\LARGE Untangling the link between traits, size and growth rate in plants

### 2016-10-19 ms/ms.tex
Polishing
Closes #46

% XX - The main comment I have is whether the strength of the 'correlation' between traits and growth rates can be summarized in some metric, so that it can be directly shown. To me Fig 4 is not very clear on showing how size affects the trait-growth correlations. For example, for wood density, I see a relation for all size classes. For the largest trees, it may look like a weaker correlation but isn't that because growth rates are lower? The same with Narea, to me they all look the same (except again, larger trees grow slower but it’s hard to tell there for me whether the trait-growth correlation is any different)
% XX -  The shade tolerance (WPLCP) analysis seems off-topic (it also not mentioned in the abstract). Can this be integrated more closely with the main objective, i.e. correlation between traits and growth rate and its dependence on size?
% XX - Georges suggests the graident of blues in Fig 4-5 not disntinct enough

@@

Here we show how a new mechanistic growth model -- called {\plant} \citep{Falster-2016} -- can explain diverse empirical phenomena, including a size-dependent effect of traits on growth and shade tolerance (Table \ref{tab:phenomena}), and thereby offer new insights into the way traits influence plant demography across the life-cycle. Broadly, the {\plant} model builds on several approaches to modelling production and allocation of biomass \citep[e.g.][]{Givnish-1988, Yokozawa-1995,Makela-1997, Moorcroft-2001, Sitch-2008, Falster-2011, King-2011}. Our primary focus in this article is to explain the pattern that has been gradually emerging -- that the effect of traits on plant growth is modified by plant size \citep{Ruger-2012, Iida-2014, Gibert-2016}. The paper by \citet{Gibert-2016} used the framework of the {\plant} model to make a conceptual argument why the effect of traits might change with size. Here we extend the results of \citet{Gibert-2016} to provide a full, functioning model and use this to how, from the point of view of primary production and allocation, why it makes sense that the effect of some traits changes with size. We also show how our approach can account for other phenomena, including changes in growth and shade tolerance with traits, individual size, and light environment. Our view is that trait-based approaches -- which aim to explain differences among species -- should be integrated within a general model of plant growth, and thus should also be able to capture patterns of growth through ontogeny. Growth rates tend to show hump-shaped relationship with size, when expressed as either height \citep{Sillett-2010, King-2011} or diameter growth \citep{Herault-2011} or biomass production \citep{Givnish-1988, Koch-2004}. In contrast, the growth rate of plant mass continues to increase with size \citep{Sillett-2010, Stephenson-2014}, and all growth measures decrease sharply with size when expressed as relative growth rates \citep{Rees-2010, Iida-2014}. Shade tolerance also varies among species, correlates with traits \citep{Messier-1999, Lusk-2008, Poorter-2006}, and tends to decrease with increasing size \citep{Givnish-1988, Kneeshaw-2006, Lusk-2008}. These diverse phenomena -- summarised in Table \ref{tab:phenomena} -- are deserving of a comprehensive and integrated explanation.

@@

Eqn \ref{eq:dbdt} can also be used to estimate a measure of shade tolerance: the light level at which a plant's photosynthetic gains just balance the costs of tissue turnover and respiration \citep{Givnish-1988, Baltzer-2007, Lusk-2013} (the ``whole-plant-light-compensation point'', {\wplcp}). In general, average assimilation rate per leaf area $\bar{p}$ increases with light level or canopy openness, $E$. The {\wplcp} can be estimated by solving for the the value of canopy openness $E=E^*$ giving $\frac{{\rm d}B} {{\rm d}t}(E)=0$ (Fig. \ref{fig:wplcp_idea}). From eqn \ref{eq:dbdt}, this occurs when
\begin{equation}\label{eq:wplcp}
\bar{p}(E^*) =\frac{\sum_{i={\rm l,b,s,r}} \, M_{\rm i} \left(\frac{k_{\rm i}}{y} + r_{\rm i}\right)}{A_{\rm l}}.
\end{equation}
The {\wplcp} occurs at the points where the photosynthetic production (per unit leaf area) line intersects with the sum of maintenance and respiration costs (per unit leaf area) for each tissue (Fig. \ref{fig:wplcp_idea}). Traits will influence the {\wplcp} if they effect either carbon uptake or costs (respiration, turnover). Also, since the amount of stem support increases with plant height, the {\wplcp} also naturally increases with height \citep{Givnish-1988} (Fig. \ref{fig:wplcp_idea}).

@@

The {\plant} adopts a model of plant construction and function that can be considered a first-order functional-balance or function-equilibrium model, similar to those implemented by \citet{Makela-1997} and \citet{Moorcroft-2001}. We could also call it ``isometric'', because the assumptions see area-based metrics scaling to the first-power of other area-based metrics, and to the square-power of length-based metrics, such as height \citep{Huxley-1932}. Table \ref{tab:allometry} provides key equations of the model (see \citealt{Falster-2016} for full derivation). In particular, we assume that as a plant grows:
\begin{description}
\item[Assumption 1:] Its height scales to the 0.5 power of its leaf area (eqn \ref{eq:ha}).
\item[Assumptions 2:] The cross-sectional area of sapwood in the stem is proportional to its leaf area (eqn \ref{eq:as}).
\item[Assumptions 3:] The cross-sectional area of bark and phloem in the stem  is proportional to its leaf area (eqn \ref{eq:ab}).
\item[Assumptions 4:] The cross-sectional area of root surface area and therefore mass is proportional to its leaf area (eqn \ref{eq:mr}).
\item[Assumption 5:] The vertical distribution of leaf within the plant's canopy, relative to the plant's height, remains constant.
\end{description}
Assumption 1 accounts for the architectural layout of the plant. Assumptions 2-4 are realisations of the pipe model \citep{Shinozaki-1964}, whereby the cross-sectional area of conducting tissues are proportional to leaf area. To describe the vertical distribution of leaf area within the canopy of an individual plant (assumption 5), we use the model of \citet{Yokozawa-1995}, which can account for a variety of canopy profiles through a single parameter $\eta_c$, varying from 0-1 (for details, see Supplementary Materials).

@@

% Importantly, we can remove these size dependent effects by altering ??? of the assumptions in Table \ref{tab:allometry} (??).
% \footnote{What are key assumptions that induce all size-related effects? My guess is  1) changing sapwood area so that volume scales isometrically with height will do it, and 2) removing reproductive allocation. Add figure in SI?}

@@

% To understand the response of growth to these traits the following derivation is helpful.

% Consider the effect of traits on the components of height growth rate (eqn \ref{eq:dhdt}). To make analysis more tractable, assume we are dealing with a plant investing all its energy in growth, i.e. $\frac{{\rm d}M_{\rm a}}{{\rm d}B}=1$. Note also that none of  the traits considered here influence $\frac{{\rm d}H} {{\rm d}A_{\rm l}}$. Therefore,
% \begin{equation} \label{eq:G2}
% \frac{{\rm d}H}{{\rm d}t} \propto \frac{{\rm d}A_{\rm l}} {{\rm d}M_{\rm a}}  \times \frac{ {\rm d}B} {{\rm d}t}.
% \end{equation}
% The correlation of traits with growth at a given height will be given by the derivative of eqn \ref{eq:G2} with respect to some trait $x$. The derivative of $\frac{{\rm d}H}{{\rm d}t}$ with respect to some trait $x$ is then \footnote{Eqn \ref{eq:G2} is a product of the form $Y(x) = W(x) \, Z(x)$, whose derivative is $\frac{\partial{Y}}{\partial{x}} = \frac{\partial{W}(x)}{\partial{x}} \, Z(x) + W(x) \, \frac{\partial{Z}(x)}{\partial{x}}$.}
% \begin{equation} \label{eq:dg1}
% \frac{\partial \left(\frac{{\rm d}H}{{\rm d}t} \right)} {\partial x} \propto
% \frac{\partial \left(\frac{{\rm d}A_{\rm l}} {{\rm d}M_{\rm a}}\right)}{\partial x}
%  \, \frac{ {\rm d}B} {{\rm d}t}
% + \frac{{\rm d}A_{\rm l}} {{\rm d}M_{\rm a}}
% \, \frac{\partial \left( \frac{ {\rm d}B} {{\rm d}t}\right)}{\partial x}.
% \end{equation}
% If this derivative is positive (negative), there will be a positive (negative) relationship between traits and growth at the height where it is evaluated. We can simplify eqn \ref{eq:dg1} further by substituting $\frac{\partial \left(\frac{{\rm d}A_{\rm l}} {{\rm d}M_{\rm a}}\right)}
% {\partial x} = -{\frac{{\rm d}A_{\rm l}} {{\rm d}M_{\rm a}}}^2
% \, \frac{\partial \left(\frac{{\rm d}M_{\rm a}} {{\rm d}A_{\rm l}}\right)
% }{\partial x}$ (from eqn \ref{eq:daldmt}). Noting also that $\frac{{\rm d}A_{\rm l}} {{\rm d}M_{\rm a}}>0$ and $\frac{{\rm d}B} {{\rm d}t} > 0$, we find that
% \begin{equation}\label{eq:dg2}
% \frac{\partial \left(\frac{{\rm d}H}{{\rm d}t} \right)} {\partial x}  \propto \frac{
% \frac{\partial \left( \frac{ {\rm d}B} {{\rm d}t}\right)}{\partial x} }
% {\frac{ {\rm d}B} {{\rm d}t}} - \frac{ \frac{\partial \left(\frac{{\rm d}M_{\rm a}} {{\rm d}A_{\rm l}}\right)
% }{\partial x}}{\frac{{\rm d}M_{\rm a}} {{\rm d}A_{\rm l}}}.
% \end{equation}

% Eqn \ref{eq:dg2} implies that there will be a negative relationship between trait $x$ and height growth if the relative change in net mass production, due to a change in trait, is less than the relative change in the total construction cost per unit leaf area, due to a change in trait. We can also solve for trait values $x$ maximising growth rate by solving for the value of $x$ that gives $\frac{\partial \left(\frac{{\rm d}H}{{\rm d}t} \right)} {\partial x}= 0$. Crucially, key elements in eqn \ref{eq:dg2} will be modified by height. With these general results in hand, let us now consider the influence of three traits on height growth rate.

\noindent{\bf Leaf nitrogen content per leaf area, $\nu$.} The response of growth rate to changes in $\nu$ is relatively straightforward: there is an optimum value of $\nu$ that maximises height growth rate in a given light environment $E$ and does not vary with height (Fig. \ref{fig:growth_light_height}). As $E$ increases from low to high, the optimal $\nu$ also increases. The invariance of the growth-trait relationship with respect to size arises as follows. The direct physiological effect of $\nu$ is to increase the maximum potential photosynthetic rate of leaves, with a cost of higher respiration rate. Both the cost and benefits of $\nu$ appear within $\frac{ {\rm d}B} {{\rm d}t}$, implying the direction of correlation between trait and growth rate depends crucially on the change in mass production per $\nu$.
%  $\frac{\partial \left( \frac{ {\rm d}B} {{\rm d}t}\right)}{\partial x}$.
From eqn \ref{eq:dbdt}, $\frac{\partial \left( \frac{ {\rm d}B} {{\rm d}t}\right)}{\partial nu} = A_{\rm l} \, (\frac{ \partial \bar{p}(E)}{\partial nu}  - \frac{ \partial r_{\rm l}}{\partial nu} )$. Since both and $\bar{p}(E)$ and $r_{\rm l}$ are expressed per unit area and independent of height, the optimal value is also independent of plant height.

@@

% Possible add some bits to suppmat?
% Substituting $x=\phi$ into eqn \ref{eq:dg2}, we can come up with a more precise expressions for some of the terms. Note that we are evaluating these derivatives at a given value of $H$ and therefore $A_{\rm l}$ (via the allometry in eqn \ref{eq:ha}). Varying $\phi$ therefore changes $M_{\rm l}$, not $A_{\rm l}$. We can then reorganise eqn \ref{eq:dbdt} as,
% \begin{equation}\label{eq:dbdt2}
% \frac{{\rm d}B}{{\rm d}t} = c_1 - c_2  \, \phi^{1 -\beta_{{\rm kl2}}} \end{equation}
% where
% $c_1 = y \, A_{\rm l} \, (\bar{p}(E) - \nu \beta_{{\rm lf4}}) - \sum_{i={\rm b,s,r}} M_{i} \left(y\, r_{i} + k{i}\right)$
% and
% $c_2 =  A_{\rm l} \, \beta_{{\rm kl1}} \, \phi_0^{\beta_{{\rm kl2}}}$ are independent of $\phi$.
% % The derivative of mass production with respect to $\phi$ is therefore
% % \begin{equation}\label{eq:dbdt3}
% % \frac{\partial \left( \frac{ {\rm d}B} {{\rm d}t}\right)}{\partial \phi}  =
% % (\beta_{{\rm kl2}}-1) \phi ^{-\beta_{{\rm kl2}}}.
% % \end{equation}
% Eqn \ref{eq:dbdt2} indicates that if $\beta_{{\rm kl2}}=1$, mass production rate $\frac{ {\rm d}B} {{\rm d}t}$ will be independent of $\phi$. In that case, lower $\phi$ will always give a growth advantage, irrespective of plant size. However,
% evidence strongly suggests $\beta_{{\rm kl2}}>1$ \citep[see Fig. \ref{fS-leaf}, after][]{Wright-2004}. In that case, $\frac{ {\rm d}B} {{\rm d}t}$ will increase with $\phi$, because of lower losses via turnover. As $\phi$ increases, the relative increase in $\frac{ {\rm d}B} {{\rm d}t}$  eventually outweighs the relative change in $\frac{{\rm d}A_{\rm l}}{{\rm d}M_{\rm a}}$, causing the direction of correlation between growth rate and trait to shift from negative to positive. Moreover, the value of $\phi$ at which this shift occurs, which is also the value of $\phi$ optimising growth rate, is higher at greater sizes.

@@

% Substituting in $x=\rho$, we can come up with a more precise expressions for some of the terms in eqn \ref{eq:dg2}. Note that we are evaluating these derivatives at a given value of $H$ and therefore $A_{\rm l}$ and thus the volume of bark and sapwood (via the allometry in eqns \ref{eq:ha}--\ref{eq:ab}). Varying $\rho$ therefore changes $M_{\rm s}$, not $A_{\rm st}$. We can then reorganise eqn \ref{eq:dbdt} as,
% \begin{equation}\label{eq:dbdt3}
% \frac{{\rm d}B}{{\rm d}t} = c_3 - c_4 \, \rho^{1-\beta_{{\rm ks2}}}
% \end{equation}
% where
% $c_3 = y \, A_{\rm l} \, (\bar{p}(E) - \sum_{i={\rm l,b,s,r}} M_{i}\, r_{i})  - sum_{i={\rm l,r}} M_{i} k{i}$
% and
% $c_4 =  (1+b) \theta \,  \eta_c \, A_{\rm l} \, H \,  \beta_{{\rm ks1}} \, \rho_0^{\beta_{{\rm ks2}}}$ are independent of $\rho$.
% Eqn \ref{eq:dbdt3} indicates that if $\beta_{{\rm ks2}}=1$, mass production rate $\frac{ {\rm d}B} {{\rm d}t}$ will be independent of $\rho$. In that case, lower $\rho$ will always give a growth advantage, irrespective of plant size. In this analysis we assume $\beta_{{\rm ks2}}=1.25$, so that mass production decreases slightly as $\rho$ decreases.

@@

\subsection{Shade tolerance}
Combining eqn \ref{eq:wplcp} with the function-balance model in Table \ref{tab:allometry} leads to the a more specific expression for calculating {\wplcp}, as the value of $E^*$ that gives
\begin{equation}\label{eq:wplcp_2}
\bar{p}(E^*) =
      \phi \left(\frac{k_{\rm l}}{y} + r_{\rm l}\right) +
      \left(\theta \, \rho \, \eta_c \, H\right)
        \left(b\left(\frac{k_{\rm b}}{y} + r_{\rm b}\right)
            + \left(\frac{k_{\rm s}}{y} + r_{\rm s}\right)\right) +
      \alpha_{\rm r1} \left(\frac{k_{\rm r}}{y} + r_{\rm r}\right).
\end{equation}
Eqn \ref{eq:wplcp_2} indicates {\wplcp} will increase approximately linearly with $H$ and potentially vary with traits $\nu$, $\phi$, and $\rho$. With some further manipulations, it is possible to show that {\wplcp} will decrease with $\phi$ if $\beta_{{\rm kl2}} >1$. Likewise {\wplcp} will decrease with $\rho$ if $\beta_{{\rm kl2}} >1$. The parameters $\beta_{{\rm ks2}}$ and $\beta_{{\rm ks2}}$ give the slope relating tissue turnover rate to $\phi$ and $\rho$, respectively. Since in this analysis, we have assumed these criteria to hold, species with low $\phi$ and low $\rho$ are predicted to be less shade tolerant (Fig. \ref{fig:wplcp}). At low $\phi$ ($\rho$), leaf (sapwood) turnover is higher and thus a greater light income is needed to offset these costs. {\wplcp} also decreases with height because as size increases, the total amount of carbon needed to offset respiratory and turnover costs in the stem also increases \citep{Givnish-1988}. In addition, {\wplcp} varies with $\nu$. At small sizes, {\wplcp} increases with $\nu$ across the band of values typically observed in real plants, i.e. high leaf nitrogen makes seedlings shade intolerant. At larger sizes, as net production declines to zero, {\wplcp} begins to increase again for very low values of $\nu$. All of these patterns match empirically observed patterns (Table \ref{tab:phenomena}).

@@

Using a model relating plant physiological function and carbon allocation to five prominent traits, we have shown how traits impact on plant growth across the life cycle. This approach extends a widely-used theoretical model for seedlings, which links mass-based growth rate to the trait leaf mass per unit leaf area \citep{Lambers-1992, Wright-2000}, to explicitly include influences of size, light environment, and other prominent traits. During the last two decades, functional traits have captured the attention of ecologists, in large part because of the ability to organise the world's plant species along standard dimensions \citep{Westoby-2002}. However, it has remained unclear how or whether prominent traits influence growth outcomes \citep{Poorter-2008, Wright-2010,Paine-2015}. Matching a growing amount of empirical evidence (Table \ref{tab:phenomena}), this study outlines when and why the direction or strength of correlation between traits and growth rate shifts with plant size. Moreover, we show that different traits and trade-offs generate different types of response. Combined with the available empirical evidence, these results demand a fundamental shift in our understanding of plant ecological strategies, away from one in which species are thought to have a fixed growth strategy throughout their life (from slow to fast growth) \citep[e.g.][]{Grime-1977, Adler-2014, Paine-2015} to one in which traits define a size-dependent growth trajectory \citep{Gibert-2016}. Moreover, we find that growth trajectories and the ranking of traits across them are also moderated by the light environment. The {\plant} model, used here, builds on and extends several related approaches, wherein emergent outcomes such as height, diameter and mass growth arise from the interaction of different tissues and traits \citep[e.g.][]{Givnish-1988, Makela-1997, Moorcroft-2001}. This approach is quite different to models derived from metabolic scaling theory (\textsc{mst}), which derive everything from a single master ``scaling'' equation for mass growth and have thus far been unable to account for size-dependent changes in the correlation between traits and growth rate \citep{Enquist-1999, Enquist-2007}. Our approach is also fundamentally different from statistically-fitted growth models \citep[e.g.][]{Herault-2011, Ruger-2012, Iida-2014} in that it predicts rather than statistically tests for trait-based effects. In this sense, our model is designed to both explained observed phenomena (Table \ref{tab:phenomena}) and also generate new hypotheses.

@@

% The only circumstances whereby traits impact on growth but not via the pathways in Fig. \ref{fig:conceptual}, would be where growth rate was entirely disconnected from the plant's carbon budget.

@@

First, our results highlight the importance of allocation decisions and turnover costs in determining growth dynamics.
% Remko thinks the following sentence is over the top but Martin de K thought it was fine.
% :) 
Much of current ecosystem research focusses on factors effecting primary production -- photosynthesis, respiration, and resultant fluxes of carbon -- with less attention devoted to allocation and turnover (\citealp{Friend-2014}; \citealp[for comparisons of models see][]{Sitch-2008, DeKauwe-2014}). Yet, for four of the five traits considered here, trait values do not influence net primary production. In fact, the analysis with $\phi$ shows that increased growth rate can occur even at a distinct cost to the plant's carbon budget. Low $\phi$ results in high leaf turnover, such that individuals with a $\phi$ have lower mass production. It is this property that makes them shade intolerant. And yet they can still achieve a growth advantage (when small), because the benefits of cheap leaf construction outweigh the costs of high leaf turnover.

@@

Second, our results demand a shift in the way plant species are sometimes described as, being of fast or slow growth. To the extent the ranking of growth rates among individuals differing in traits shifts with either plant size or light environment, it is not possible to describe a species via a single point along a spectrum from slow to fast growth. Such a spectrum is implied by many of theoretical models used in community ecology, including Grime's \textsc{csr} triangle, the r-K spectrum, and coexistence models base on the Lotka-Volterra system of equations \citep[e.g.][]{Grime-1977,Chesson-2000}. Researchers using functional traits have also tended to describe species as fast or slow growing \citep[e.g.][]{Adler-2014, Diaz-2016}. Our results suggest a more-nuanced approach. Plants that are fast growing as seedlings may not be fast growing as saplings or adults, or under low light. Plants that are fast growing as adults may not be fast growing as seedlings. This more-nuanced perspective tends to mirror observed demographic patterns, where juvenile and adult growth rates are sometimes only loosely correlated \citep{Rees-2001}.
% So we should instead be thinking of a species traits as defining a trajectory of potential growth for an individual, and one that is moderated by plant size and light environment.

@@

Another class of model dealing explicitly with size-related effects includes those derived from the metabolic scaling theory (\textsc{mst}) of ecology \citep{Enquist-1999, Enquist-2007}. Several points suggest our new framework provides a better explanation to the growth phenomena in Table \ref{tab:phenomena} than the \textsc{mst} framework. First, the \textsc{mst}-derived model suggests diameter growth continues to increase as plants grow, whereas empirical data suggests growth rate declines for larger plants \citep{Herault-2011}. Second, the \textsc{mst} model does not allow for the effects of traits to vary with plant size. Predicted effects are for a linear increase in growth with decreases in either $\phi$ and lower $\rho$, that apply irrespective of size. However, at least for $\phi$ such effects in large trees have not been observed.
% yes, I like this shortened version!

@@

\section{Tables and Figures}

>>>>

@@

Here we show how a new mechanistic growth model -- called {\plant} \citep{Falster-2016} -- can explain diverse empirical phenomena, including a size-dependent effect of traits on growth and an effect of traits on shade tolerance (Table \ref{tab:phenomena}), and thereby offer new insights into the way traits influence plant demography across the life-cycle. Broadly, the {\plant} model builds on several approaches to modelling production and allocation of biomass \citep[e.g.][]{Givnish-1988, Yokozawa-1995,Makela-1997, Moorcroft-2001, Sitch-2008, Falster-2011, King-2011}. Our primary focus in this article is to explain the pattern that has been gradually emerging -- that the effect of traits on plant growth is modified by plant size \citep{Ruger-2012, Iida-2014, Gibert-2016}. Based on the same decomposition of growth rates as is implemented in the {\plant} model and used below \citep[from][]{Falster-2011}, \citet{Gibert-2016} argued conceptually why the effect of traits on growth should change with size. Here we extend the results of \citet{Gibert-2016} to provide a full, functioning model and use this to show, from the point of view of primary production and allocation, how and why the effect of some traits on growth rates changes with size. We also show how our approach can account for other phenomena, including changes in growth and shade tolerance with traits, individual size, and light environment. Our view is that trait-based approaches -- which aim to explain differences among species -- should be integrated within a general model of plant growth, and thus should also be able to capture patterns of growth through ontogeny. Growth rates tend to show hump-shaped relationship with size, when expressed as either height \citep{Sillett-2010, King-2011} or diameter growth \citep{Canham-2004, Canham-2006, Herault-2011} or biomass production \citep{Givnish-1988, Koch-2004}. In contrast, the growth rate of standing plant mass continues to increase with size \citep{Sillett-2010, Stephenson-2014}. All growth measures decrease sharply with size when expressed as relative growth rates \citep{Rees-2010, Iida-2014}. Shade tolerance also varies among species, correlates with traits \citep{Messier-1999, Lusk-2008, Poorter-2006}, and tends to decrease with increasing size \citep{Givnish-1988, Kneeshaw-2006, Lusk-2008}. These diverse phenomena -- summarised in Table \ref{tab:phenomena} -- are deserving of a comprehensive and integrated explanation.

@@

Eqn \ref{eq:dbdt} can also be used to estimate a measure of shade tolerance: the light level at which a plant's photosynthetic gains just balance the costs of tissue turnover and respiration \citep{Givnish-1988, Baltzer-2007, Lusk-2013} (the ``whole-plant-light-compensation point'', {\wplcp}). In general, average assimilation rate per leaf area $\bar{p}$ increases with light level or canopy openness, $E$. The {\wplcp} can be estimated by solving for the the value of canopy openness $E=E^*$ giving $\frac{{\rm d}B} {{\rm d}t}(E)=0$ (Fig. \ref{fig:wplcp_idea}). From eqn \ref{eq:dbdt}, this occurs when
\begin{equation}\label{eq:wplcp}
\bar{p}(E^*) =\frac{\sum_{i={\rm l,b,s,r}} \, M_{\rm i} \left(\frac{k_{\rm i}}{y} + r_{\rm i}\right)}{A_{\rm l}}.
\end{equation}
The {\wplcp} occurs at the points where the photosynthetic production (per unit leaf area) line intersects with the sum of maintenance and respiration costs (per unit leaf area) for each tissue (Fig. \ref{fig:wplcp_idea}). Traits can influence the {\wplcp} if they effect either carbon uptake or costs (respiration, turnover). Also, since the amount of stem support increases with plant height, the {\wplcp} also naturally increases with height \citep{Givnish-1988} (Fig. \ref{fig:wplcp_idea}).

@@

The {\plant} package adopts a model of plant construction and function that can be considered a first-order functional-balance or function-equilibrium model, similar to those implemented by \citet{Makela-1997} and \citet{Moorcroft-2001}. We could also call it ``isometric'', because the assumptions see area-based metrics scaling to the first-power of other area-based metrics, and to the square-power of length-based metrics, such as height \citep{Huxley-1932}. Table \ref{tab:allometry} provides key equations of the model (see \citealt{Falster-2016} for full derivation). In particular, we assume that as a plant grows:
\begin{description}
\item[Assumption 1:] Its height scales to the 0.5 power of its leaf area (eqn \ref{eq:ha}).
\item[Assumptions 2:] The cross-sectional area of sapwood in the stem is proportional to its leaf area (eqn \ref{eq:as}).
\item[Assumptions 3:] The cross-sectional area of bark and phloem in the stem  is proportional to its leaf area (eqn \ref{eq:ab}).
\item[Assumptions 4:] The cross-sectional area of root surface area and therefore mass is proportional to its leaf area (eqn \ref{eq:mr}).
\item[Assumption 5:] The vertical distribution of leaf within the plant's canopy, relative to the plant's height, remains constant.
\end{description}
Assumption 1 accounts for the architectural layout of the plant. Assumptions 2-4 are realisations of the pipe model \citep{Shinozaki-1964}, whereby the cross-sectional area of conducting tissues are proportional to leaf area. To describe the vertical distribution of leaf area within the canopy of an individual plant (assumption 5), we use the model of \citet{Yokozawa-1995}, which can account for a variety of canopy profiles through a single parameter $\eta_c$, varying from 0-1 (for details, see Supplementary Materials).

@@

@@

\noindent{\bf Leaf nitrogen content per leaf area, $\nu$.} The response of growth rate to changes in $\nu$ is relatively straightforward: there is an optimum value of $\nu$ that maximises height growth rate in a given light environment $E$ and does not vary with height (Fig. \ref{fig:growth_light_height}). As $E$ increases from low to high, the optimal $\nu$ also increases. The invariance of the growth-trait relationship with respect to size arises as follows. The direct physiological effect of $\nu$ is to increase the maximum potential photosynthetic rate of leaves, with a cost of higher respiration rate. Both the cost and benefits of $\nu$ appear within $\frac{ {\rm d}B} {{\rm d}t}$, implying the direction of correlation between trait and growth rate depends crucially on the change in mass production per $\nu$.
From eqn \ref{eq:dbdt}, $\frac{\partial \left( \frac{ {\rm d}B} {{\rm d}t}\right)}{\partial nu} = A_{\rm l} \, (\frac{ \partial \bar{p}(E)}{\partial nu}  - \frac{ \partial r_{\rm l}}{\partial nu} )$. Since both and $\bar{p}(E)$ and $r_{\rm l}$ are expressed per unit area and independent of height, the optimal value is also independent of plant height.

@@

@@

@@

\subsection{Shade tolerance}
Combining eqn \ref{eq:wplcp} with the function-balance model in Table \ref{tab:allometry} leads to the a more specific expression for calculating {\wplcp}, as the value of $E^*$ that gives
\begin{equation}\label{eq:wplcp_2}
\bar{p}(E^*) =
      \phi \left(\frac{k_{\rm l}}{y} + r_{\rm l}\right) +
      \left(\theta \, \rho \, \eta_c \, H\right)
        \left(b\left(\frac{k_{\rm b}}{y} + r_{\rm b}\right)
            + \left(\frac{k_{\rm s}}{y} + r_{\rm s}\right)\right) +
      \alpha_{\rm r1} \left(\frac{k_{\rm r}}{y} + r_{\rm r}\right).
\end{equation}
Eqn \ref{eq:wplcp_2} indicates {\wplcp} will increase approximately linearly with $H$ and potentially vary with traits $\nu$, $\phi$, and $\rho$. With some further manipulations, it is possible to show that {\wplcp} will decrease with $\phi$ if $\beta_{{\rm kl2}} >1$. Likewise {\wplcp} will decrease with $\rho$ if $\beta_{{\rm ks2}} >1$. The parameters $\beta_{{\rm kl2}}$ and $\beta_{{\rm ks2}}$ give the slope relating tissue turnover rate to $\phi$ and $\rho$, respectively. Since in this analysis, we have assumed these criteria to hold, species with low $\phi$ and low $\rho$ are predicted to be less shade tolerant because of disproportionate increases in turnover costs (Fig. \ref{fig:wplcp}). At low $\phi$ ($\rho$), leaf (sapwood) turnover is higher and thus a greater light income is needed to offset these costs. {\wplcp} also decreases with height because as size increases, the total amount of carbon needed to offset respiratory and turnover costs in the stem also increases \citep{Givnish-1988}. In addition, {\wplcp} varies with $\nu$. At small sizes, {\wplcp} increases with $\nu$ across the band of values typically observed in real plants, i.e. high leaf nitrogen makes seedlings shade intolerant. At larger sizes, as net production declines to zero, {\wplcp} begins to increase again for very low values of $\nu$. All of these patterns match empirically observed patterns (Table \ref{tab:phenomena}).

@@

Using a model relating plant physiological function and carbon allocation to five prominent traits, we have shown how traits impact on plant growth across the life cycle. This approach extends a widely-used theoretical model for seedlings, which links mass-based growth rate to the trait leaf mass per unit leaf area \citep{Lambers-1992, Wright-2000}, to explicitly include influences of size, light environment, and other prominent traits. During the last two decades, functional traits have captured the attention of ecologists, in large part because of the ability to organise the world's plant species along standard dimensions \citep{Westoby-2002}. However, it has remained unclear how or whether prominent traits influence growth outcomes \citep{Poorter-2008, Wright-2010,Paine-2015}. Matching a growing amount of empirical evidence (Table \ref{tab:phenomena}), this study outlines when and why the direction or strength of correlation between traits and growth rate shifts with plant size. Moreover, we show that different traits and trade-offs generate different types of response. Combined with the available empirical evidence, these results demand a fundamental shift in our understanding of plant ecological strategies, away from one in which species are thought to have a fixed growth strategy throughout their life (from slow to fast growth) \citep[e.g.][]{Grime-1977, Adler-2014, Paine-2015} to one in which traits define a size-dependent growth trajectory \citep{Gibert-2016}. Moreover, we find that growth trajectories and the ranking of traits across them are also moderated by the light environment; while traits that minimise costs of tissue respiration and / or turnover also make plants more shade tolerant (i.e. lower \textsc{wplcp}), as is empirically observed \citep{Messier-1999, Craine-2005, Poorter-2006, Baltzer-2007, Lusk-2008}. The {\plant} model, used here, builds on and extends several related approaches, wherein emergent outcomes such as height, diameter and mass growth arise from the interaction of different tissues and traits \citep[e.g.][]{Givnish-1988, Makela-1997, Moorcroft-2001}. This approach is quite different to models derived from metabolic scaling theory (\textsc{mst}), which derive everything from a single master ``scaling'' equation for mass growth and have thus far been unable to account for size-dependent changes in the correlation between traits and growth rate \citep{Enquist-1999, Enquist-2007}. Our approach is also fundamentally different from statistically-fitted growth models \citep[e.g.][]{Herault-2011, Ruger-2012, Iida-2014} in that it predicts rather than statistically tests for trait-based effects. In this sense, our model is designed to both explained observed phenomena (Table \ref{tab:phenomena}) and also generate new hypotheses.

@@

@@

First, our results highlight the importance of allocation decisions and turnover costs in determining growth dynamics.
Much of current ecosystem research focusses on factors effecting primary production -- photosynthesis, respiration, and resultant fluxes of carbon -- with less attention devoted to allocation and turnover (\citealp{Friend-2014}; \citealp[for comparisons of models see][]{Sitch-2008, DeKauwe-2014}). Yet, for four of the five traits considered here, trait values do not influence net primary production. In fact, the analysis with $\phi$ shows that increased growth rate can occur even at a distinct cost to the plant's carbon budget. Low $\phi$ results in high leaf turnover, such that individuals with a $\phi$ have lower mass production. It is this property that makes them shade intolerant. And yet they can still achieve a growth advantage (when small), because the benefits of cheap leaf construction outweigh the costs of high leaf turnover.

@@

Second, our results demand a shift in the way plant species are sometimes described as, being of fast or slow growth. To the extent the ranking of growth rates among individuals differing in traits shifts with either plant size or light environment, it is not possible to describe a species via a single point along a spectrum from slow to fast growth. Such a spectrum is implied by many of theoretical models used in community ecology, including Grime's \textsc{csr} triangle, the r-K spectrum, and coexistence models base on the Lotka-Volterra system of equations \citep[e.g.][]{Grime-1977,Chesson-2000}. Researchers using functional traits have also tended to describe species as fast or slow growing \citep[e.g.][]{Adler-2014, Diaz-2016}. Our results suggest a more-nuanced approach. Plants that are fast growing as seedlings may not be fast growing as saplings or adults, or under low light. Plants that are fast growing as adults may not be fast growing as seedlings. This more-nuanced perspective tends to mirror observed demographic patterns, where juvenile and adult growth rates are sometimes only loosely correlated \citep{Rees-2001}.

@@

Another class of model dealing explicitly with size-related effects includes those derived from the metabolic scaling theory (\textsc{mst}) of ecology \citep{Enquist-1999, Enquist-2007}. Several points suggest our new framework provides a better explanation to the growth phenomena in Table \ref{tab:phenomena} than the \textsc{mst} framework. First, the \textsc{mst}-derived model suggests diameter growth continues to increase as plants grow, whereas empirical data suggests growth rate declines for larger plants \citep{Canham-2004, Canham-2006, Herault-2011}. Second, the \textsc{mst} model does not allow for the effects of traits to vary with plant size. Predicted effects are for a linear increase in growth with decreases in either $\phi$ and lower $\rho$, that apply irrespective of size. However, at least for $\phi$ such effects in large trees have not been observed.

@@

\newpage
\section{Tables and Figures}
